Results

### 2017-11-17 tmp-results.Rmd
tmp results to discuss

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.asp = 0.7)
options(width = 110)
```

@@

@@

```{r, fig.width=16, out.width="95%", dev="svg"}
ttrees_vs_logit <- filter(results, method %in% c("T-Trees", "logit-simple"))

@@

```{r, fig.width=14, out.width="90%", dev="svg"}
H2 <- 0.8
cowplot::plot_grid(
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases10, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 10%", 
         color = "Distribution\nof effects"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases20, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 20%", 
         color = "Distribution\nof effects"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases10, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 10%", 
         color = "Model"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases20, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 20%", 
         color = "Model"),
  labels = LETTERS[1:4], align = "hv", label_size = 25, scale = 0.95
)
```

@@

```{r}
# represent h2=0.8 as a function of h2=0.5
results2 %>%
  select(starts_with("par"), method, AUC) %>%
  group_by(par.causal, par.dist, par.model, method, par.h2) %>%
  summarise(AUC = mean(AUC)) %>%
  spread(par.h2, AUC) %>%
  myggplot(aes(`0.5`, `0.8`, color = method)) +
  geom_smooth(size = 2, alpha = 0.2) +
  geom_point(size = 3)

@@

```{r, fig.cap="Main results: logit simple vs PRS max"}
results2 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         method %in% c("logit-simple", "PRS-max")) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC, 1e4)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

```{r, fig.cap="Main results: all PRS"}
results2 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         grepl("PRS", method)) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC),
            AUC_boot = boot(AUC, 1e4)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

@@

```{r, fig.cap="Main results: logit simple vs PRS max for chromosome 6"}
results3 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         method %in% c("logit-simple", "PRS-max")) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC, 1e4)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

```{r, fig.cap="Main results: all PRS for chromosome 6"}
results3 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         grepl("PRS", method)) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC),
            AUC_boot = boot(AUC, 1e4)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

@@

```{r}
knitr::include_graphics("figures/celiac-man.png")
```

```{r}
knitr::include_graphics("figures/celiac-regpath.png")
```

@@

```{r}
results2 %>%
  filter(
    # method %in% c("logit-simple", "PRS-max"),
    grepl("all", par.causal)
  ) %>%
  mutate(M = readr::parse_number(par.causal)) %>%
  group_by(M, par.dist, par.h2, par.model, method) %>%
  summarise(AUC = mean(AUC)) %>%
  # filter(par.h2 == 0.8, par.dist == "gaussian", par.model == "simple") %>%
  filter(par.h2 == 0.8) %>%
  myggplot(aes(M, AUC, color = method)) +
  geom_line() + 
  geom_point(size = 2) +
  facet_grid(par.dist ~ par.model) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10) +
  scale_x_sqrt(breaks = c(30, 300, 1200, 3000)) +
  labs(x = "Number of causal SNPs (sqrt-scale)")
```

```{r}
results3 %>%
  filter(
    # method %in% c("logit-simple", "PRS-max"),
    grepl("all", par.causal)
  ) %>%
  mutate(M = readr::parse_number(par.causal)) %>%
  group_by(M, par.dist, par.h2, par.model, method) %>%
  summarise(AUC = mean(AUC)) %>%
  # filter(par.h2 == 0.8, par.dist == "gaussian", par.model == "simple") %>%
  filter(par.h2 == 0.8) %>%
  myggplot(aes(M, AUC, color = method)) +
  geom_line() + 
  geom_point(size = 2) +
  facet_grid(par.dist ~ par.model) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10) +
  scale_x_sqrt(breaks = c(30, 300, 1200, 3000)) +
  labs(x = "Number of causal SNPs (sqrt-scale)")
```

```{r}
bind_rows(
  bind_cols(results2, simu = rep("all", nrow(results2))),
  bind_cols(results3, simu = rep("chr6", nrow(results3)))
) %>%
  filter(par.h2 == 0.8, par.dist == "gaussian", method != "logit-triple") %>%
  myggplot() +
  geom_boxplot(aes(method, AUC, fill = simu, color = simu), alpha = 0.3) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(par.model ~ par.causal) +
  theme(strip.text.x = element_text(size = rel(2)),
        strip.text.y = element_text(size = rel(2))) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  geom_hline(yintercept = 0.94, color = "blue", linetype = 2)
```

```{r, fig.cap="Size of effects GWAS vs logistic"}
knitr::include_graphics("effects.png")
```

```{r}
knitr::include_graphics("preds-density2.png")
```

```{r}
knitr::include_graphics("preds-density3.png")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.asp = 0.7,
                      dev = "svg")
options(width = 110)
```

@@

```{r}
boot <- function(x, n = 1e5, f = mean) {
  sd(replicate(n, f(sample(x, replace = TRUE))))
}
```

@@

```{r, fig.width=16, out.width="95%", dev="svg", fig.cap="Results of T-Trees vs penalized logistic regression. **A.** Timing (in seconds). **B.** Number of predictors of the model. **C.** AUC. **D.** Percentage of cases in the 10% largest scores."}
ttrees_vs_logit <- filter(results, method %in% c("T-Trees", "logit-simple"))

@@

```{r, fig.width=14, out.width="90%", dev="svg", fig.cap="Percentage of cases in the 2 highest deciles of PRSs as a function of AUC."}
H2 <- 0.8
cowplot::plot_grid(
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases10, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 10%", 
         color = "Distribution\nof effects"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases20, color = par.dist)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    labs(y = "Percentage of cases in top 20%", 
         color = "Distribution\nof effects"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases10, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 10%", 
         color = "Model"),
  results2 %>%
    filter(par.h2 == H2) %>%
    myggplot(aes(AUC, percCases20, color = par.model)) +
    geom_point() +
    geom_smooth(method = "lm") +
    theme(legend.position = c(0.8, 0.2)) +
    scale_colour_brewer(palette = "Set1") +
    labs(y = "Percentage of cases in top 20%", 
         color = "Model"),
  labels = LETTERS[1:4], align = "hv", label_size = 25, scale = 0.95
)
```

@@

```{r, fig.cap="Results of AUC for all combination of parameters and methods when h2=0.5 and h2=0.8"}
# represent h2=0.8 as a function of h2=0.5
results2 %>%
  select(starts_with("par"), method, AUC) %>%
  group_by(par.causal, par.dist, par.model, method, par.h2) %>%
  summarise(AUC = mean(AUC)) %>%
  spread(par.h2, AUC) %>%
  myggplot(aes(`0.5`, `0.8`, color = method)) +
  geom_smooth(size = 2, alpha = 0.2) +
  geom_point(size = 3)

@@

```{r, fig.cap="Main results: logit simple vs PRS max"}
results2 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         method %in% c("logit-simple", "PRS-max")) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC, 1e5, mean)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

```{r, fig.cap="Main results: all PRS"}
results2 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         grepl("PRS", method)) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC),
            AUC_boot = boot(AUC, 1e5, mean)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

@@

```{r, fig.cap="Main results: logit simple vs PRS max for chromosome 6"}
results3 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         method %in% c("logit-simple", "PRS-max")) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC), AUC_boot = boot(AUC, 1e5, mean)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

```{r, fig.cap="Main results: all PRS for chromosome 6"}
results3 %>%
  filter(par.dist == "laplace", par.h2 == 0.8, par.model == "simple",
         grepl("PRS", method)) %>%
  group_by(par.causal, method) %>%
  summarise(AUC_mean = mean(AUC),
            AUC_boot = boot(AUC, 1e5, mean)) %>%
  myggplot(aes(par.causal, AUC_mean, fill = method, color = method)) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_bar(stat = "identity", alpha = 0.3, position=position_dodge()) +
  geom_errorbar(aes(ymin = AUC_mean - 2 * AUC_boot, ymax = AUC_mean + 2 * AUC_boot),
                position=position_dodge(width=0.9), color = "black", width = 0.2) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = 0:10 / 10 + 0.05)
```

@@

```{r, fig.cap="Manhanttan plot for Celiac"}
knitr::include_graphics("figures/celiac-man.png")
```

```{r, fig.cap="Regularization paths for the methods. For LR, line in the result given by CMSA."}
knitr::include_graphics("figures/celiac-regpath.png")
```

@@

```{r, fig.cap="AUC as function of M. For all chromosomes."}
results2 %>%
  filter(
    # method %in% c("logit-simple", "PRS-max"),
    grepl("all", par.causal)
  ) %>%
  mutate(M = readr::parse_number(par.causal)) %>%
  group_by(M, par.dist, par.h2, par.model, method) %>%
  summarise(AUC = mean(AUC)) %>%
  # filter(par.h2 == 0.8, par.dist == "gaussian", par.model == "simple") %>%
  filter(par.h2 == 0.8) %>%
  myggplot(aes(M, AUC, color = method)) +
  geom_line() + 
  geom_point(size = 2) +
  facet_grid(par.dist ~ par.model) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10) +
  scale_x_sqrt(breaks = c(30, 300, 1200, 3000)) +
  labs(x = "Number of causal SNPs (sqrt-scale)")
```

```{r, fig.cap="AUC as function of M. For chromosome 6."}
results3 %>%
  filter(
    # method %in% c("logit-simple", "PRS-max"),
    grepl("all", par.causal)
  ) %>%
  mutate(M = readr::parse_number(par.causal)) %>%
  group_by(M, par.dist, par.h2, par.model, method) %>%
  summarise(AUC = mean(AUC)) %>%
  # filter(par.h2 == 0.8, par.dist == "gaussian", par.model == "simple") %>%
  filter(par.h2 == 0.8) %>%
  myggplot(aes(M, AUC, color = method)) +
  geom_line() + 
  geom_point(size = 2) +
  facet_grid(par.dist ~ par.model) +
  scale_y_continuous(breaks = 0:10 / 10, minor_breaks = c(0:9 + 0.5) / 10) +
  scale_x_sqrt(breaks = c(30, 300, 1200, 3000)) +
  labs(x = "Number of causal SNPs (sqrt-scale)")
```

```{r, fig.cap="AUCs for h2=0.8, dist=gaussian, comparing all chromosomes and chromosome 6."}
bind_rows(
  bind_cols(results2, simu = rep("all", nrow(results2))),
  bind_cols(results3, simu = rep("chr6", nrow(results3)))
) %>%
  filter(par.h2 == 0.8, par.dist == "gaussian", method != "logit-triple") %>%
  myggplot() +
  geom_boxplot(aes(method, AUC, fill = simu, color = simu), alpha = 0.3) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(par.model ~ par.causal) +
  theme(strip.text.x = element_text(size = rel(2)),
        strip.text.y = element_text(size = rel(2))) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  geom_hline(yintercept = 0.94, color = "blue", linetype = 2)
```

```{r, fig.cap="Size of effects GWAS vs logistic, for Celiac"}
knitr::include_graphics("effects.png")
```

```{r, fig.cap="Density of scores from logistic regression by pop."}
knitr::include_graphics("preds-density2.png")
```

```{r, fig.cap="Density of scores from logistic regression by pop and genotype."}
knitr::include_graphics("preds-density3.png")
```
